Supporting information: Tail associations and their impact on extinction risk

### 2019-11-09 SuppMat.Rmd
fixed typos

<!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
Bivarite environmental noise with extreme left-tail association (as in Fig. 
\ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
by @Ghosh_copula (their Appendix S10), presented again here for convenience.
If $T$ time steps of noise were desired, step 1 was to generate a $T$ by $2$
matrix, $M$, of independent random numbers, each standard-normally distributed. 
Then, each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association (as in Fig. 
\ref{MT-fig_intro_cop}D), take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed.
The Spearman correlation of the extreme left- or right-tail associated 
noise was `r srho_common`, up to sampling variation. A derivation
of this fact is presented below.

<!--moderate tail association noise-->
Bivarite environmental noise with moderate left-tail association (as in Fig. \ref{MT-fig_intro_cop}B) was 
generated using a 2-dimensional Clayton copula. The statistical
theory of copulas has been introduced systematically in several texts 
(e.g., @nelsen2006_copula; @joe2014_dependence), and was briefly 
introduced by @Genest2007. Short, conceptual introductions aimed at ecological
audiences were provided by sections within the papers of @Anderson2018 and @Ghosh_copula. A
2-dimensional Clayton copula is equivalent to a particular bivariate 
probability density function with marginal distributions which are 
uniform on the interval $(0,1)$. The points on Fig. 
\ref{MT-fig_intro_cop}B were independent draws from the associated random variable.
The functions `claytonCopula` and `rCopula` from the `copula` package for the 
R programming language were used. The parameter of the Clayton family was
chosen so that the Spearman correlation was `r srho_common`, the same as that of the
extreme left- and right-tail associated noise. This was accomplished with the `iRho`
function of the `copula` package. Noise with moderate right-tail
association was generated in a similar way but using the 2-dimensional 
survival Clayton copula. The survival Clayton copula is a rotation
of the Clayton copula by 180 degrees.

<!--Multivariate extreme tail association-->
Multivariate noise with extreme left-tail association was generated using a
very simlar algorithm to the bivariate case,
as follows. If $T$ time steps of noise of dimension $n$ were desired,
step 1 was to generate a $T$ by $n$
matrix, $M$, of independent random numbers, each standard-normally distributed.
Each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association, take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed. 
Multivariate moderately tail associated noise of dimension $n$
was generated using $n$-dimensional Clayton or survival Clayton copulas.

>>>>

<!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
Bivariate environmental noise with extreme left-tail association (as in Fig. 
\ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
by @Ghosh_copula (their Appendix S10), presented again here for convenience.
If $T$ time steps of noise were desired, step 1 was to generate a $T$ by $2$
matrix, $M$, of independent random numbers, each standard-normally distributed. 
Then, each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association (as in Fig. 
\ref{MT-fig_intro_cop}D), take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed.
The Spearman correlation of the extreme left- or right-tail associated 
noise was `r srho_common`, up to sampling variation. A derivation
of this fact is presented below.

<!--moderate tail association noise-->
Bivariate environmental noise with moderate left-tail association (as in Fig. \ref{MT-fig_intro_cop}B) was 
generated using a 2-dimensional Clayton copula. The statistical
theory of copulas has been introduced systematically in several texts 
(e.g., @nelsen2006_copula; @joe2014_dependence), and was briefly 
introduced by @Genest2007. Short, conceptual introductions aimed at ecological
audiences were provided by sections within the papers of @Anderson2018 and @Ghosh_copula. A
2-dimensional Clayton copula is equivalent to a particular bivariate 
probability density function with marginal distributions which are 
uniform on the interval $(0,1)$. The points in Fig. 
\ref{MT-fig_intro_cop}B were independent draws from the associated random variable.
The functions `claytonCopula` and `rCopula` from the `copula` package for the 
R programming language were used. The parameter of the Clayton family was
chosen so that the Spearman correlation was `r srho_common`, the same as that of the
extreme left- and right-tail associated noise. This was accomplished with the `iRho`
function of the `copula` package. Noise with moderate right-tail
association was generated in a similar way but using the 2-dimensional 
survival Clayton copula. The survival Clayton copula is a rotation
of the Clayton copula by 180 degrees.

<!--Multivariate extreme tail association-->
Multivariate noise with extreme left-tail association was generated using a
very similar algorithm to the bivariate case,
as follows. If $T$ time steps of noise of dimension $n$ were desired,
step 1 was to generate a $T$ by $n$
matrix, $M$, of independent random numbers, each standard-normally distributed.
Each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association, take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed. 
Multivariate moderately tail associated noise of dimension $n$
was generated using $n$-dimensional Clayton or survival Clayton copulas.
